---
title: "p8105_hw3_fz2328"
author: "Fengdi Zhang"
date: "2022-10-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(ggridges)
library(patchwork)
```

## Problem 2

Load, tidy and wrangle the data.

```{r, message=FALSE}
accel_data = 
      read_csv("dataset/accel_data.csv") %>% 
      janitor::clean_names() %>%
      pivot_longer(
              activity_1:activity_1440,
              names_to = "minute_day",
              names_prefix = "activity_",
              values_to = "activity_counts") %>% 
      mutate(
        week_factor = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"),
        minute_day = as.numeric(minute_day)
      )
 
```

The resulting dataset includes `r ncol(accel_data)` variables, including `r ls(accel_data)`. There are `r nrow(accel_data)` observations in total. 


```{r, message=FALSE}
accel_data %>%
  group_by(week, day) %>%
  mutate(activity_day_total = sum(activity_counts)) %>% 
  group_by(week, day, activity_day_total) %>% 
  summarize() %>% 
  pivot_wider(
    names_from = "day",
    values_from = "activity_day_total") %>% 
  select(week, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable(digits = 1)
```

It seems that the activity counts are much lower in the first 2 days in the first week and during the weekends in the last 2 weeks. Otherwise, there is no obvious trend.

```{r}
accel_data %>% 
  group_by(week, day) %>% 
  ggplot(aes(x = minute_day, y = activity_counts, color = day)) + 
  geom_point(alpha = .5 ) 
  
```

According to the graph, wee can see that this patient tends to have more activities at certain times during the days, such as 650mins/10:50am, 1000mins/4:40pm, and 1250mins/8:50pm. 


## Problem 3

First, import the data
```{r, message=FALSE}
library(p8105.datasets)
data("ny_noaa")
```

There are `r nrow(accel_data)` observations in total. The dataset includes `r ncol(ny_noaa)` variables, including `r ls(accel_data)`. There are missing values in these variables: `r names(which(colSums(is.na(ny_noaa))>0))`. The missing values would cause problems when summarizing the variables statistics, restructuring variables, and so on. 

Let's tidy the data first
```{r}
ny_noaa = ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-")

ny_noaa = ny_noaa %>% 
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day),
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin)
  ) %>% 
  mutate(
    prcp = 0.1*prcp, 
    tmax = 0.1*tmax,
    tmin = 0.1*tmax
  )

```

Now, year, month and day are separate variables. Precipitation, snowfall and snow depth have a unit of mm. Temperatures have a unit of degrees C. 

The most commonly observed values for snowfall is `r names(which.max(table(ny_noaa$snow)))`. This is because New York only have a few days of snowing each year. The other reason can be that there were about one half of the stations report precipitation only. 


Now, let's make a two-panel plot showing the average max temperature in January and in July in each station across years.
```{r, message=FALSE, warning=FALSE}
ny_noaa %>% 
  filter(month == 1 | month == 7) %>% 
  group_by(id, year, month) %>% 
  mutate(tmax_mean = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = tmax_mean, group = id)) + 
  geom_line(alpha = .5) + 
  facet_grid(. ~ month)
```

In general, the average max temperature in July is significantly higher than the average max temperature in January. 
The max temperature in July has a narrower range than the max temperature in January. 
In January, there are a bunch of outliers above and below the average of tmax. In July, most outliers lie below the average of tmax. 
Each station has similar temperature change patterns. 


Now, let's make a two-panel plot showing (i) `tmax` vs `tmin` for the full dataset; (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
```{r, message=FALSE, warning=FALSE}
A = ny_noaa %>% 
  ggplot(aes(x = tmin, y = tmax, color = month)) + 
  geom_hex()
 
B = ny_noaa %>%
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, fill = year, group = year)) + 
  geom_density(alpha = .4, adjust = .5) 

A/B
  
```

